Draft_Application_Stochastic_Volatility

### 2021-07-16 Real Dataset SV.Rmd
Aggiunte

\section{Overview}
In the following sections, we will try to replicate some of the analyses which were previously conducted on synthetic data, generated by a basic stochastic volatility model; this time, the same model specification and filters will be used on real data. However, as we will see, real data require different strategies to evaluate the effectiveness of several Sequential Monte Carlo techniques. Intuitively, such differences are due to the fact that now the latent process driving the Hidden Markov Model is not observed, nor do we know exactly its law of motion. 
This part of the work is structured as follows: section \textit{fill with the number of the section} provides a review of recent research in the stochastic volatility literature, including extensions of the basic specification and the use of particle filters; section \textit{fill with the number of the section} describes the data which we employ and the transformations that were applied to them. In section \textit{fill with the number of the section} we briefly compare the series of filtered states that are obtained from different particle filters; moreover, several PFs are compared in terms of computational cost in approximating the exact filtering distribution implied by our SV model. Section \textit{fill with the number of the section} features a forecast comparison between two SV specifications and a model that assumes constant mean and variance for the financial returns, namely a forecasting exercise that would not be possible were it not for particle filtering techniques. In section \textit{fill with the number of the section} we conclude by employing a non-parametric measure of realised volatility as a proxy for the latent volatility process, allowing a RMSE and MAE comparison between filters which is similar to that of section \textit{fill with the number of the section} for synthetic data.

@@

As regards the series that we will employ in section \textit{fill with the number of the section}, namely the series of realised volatilities which we will employ as a proxy for the true latent volatilities, the data were retrieved from the Oxford-Man Institute's Realized Library, which provides several different daily non-parametric measures of past volatility. We consider a proper rescaling of the \textit{rv5} series, \textit{i.e.} model-free daily volatility estimates based on 5 min intraday return intervals. Also in this case, the chosen interval spanned from June 1st, 2017 to May 30th, 2021. \textit{Includere grafico e commento del grafico della realised volatility}

\subsection{Transformations}
Inserire le trasformazioni effettuate (e il perchè)

>>>>

\section{Overview}

In the following sections, we will try to replicate some of the analyses which were previously conducted on synthetic data, generated by a basic stochastic volatility model; this time, the same model specification and filters will be used on real data. However, as we will see, real data require different strategies to evaluate the effectiveness of several Sequential Monte Carlo techniques. Intuitively, such differences are due to the fact that now the latent process driving the Hidden Markov Model is not observed, nor do we know exactly its law of motion. 
This part of the work is structured as follows: section \textit{fill with the number of the section} provides a review of recent research in the stochastic volatility literature, including extensions of the basic specification and the use of particle filters; section \textit{fill with the number of the section} describes the data which we employ and the transformations that were applied to them. In section \textit{fill with the number of the section} we briefly compare the series of filtered states that are obtained from different particle filters; moreover, several PFs are compared in terms of computational cost in approximating the exact filtering distribution implied by our SV model. Section \textit{fill with the number of the section} features a forecast comparison between two SV specifications and a model that assumes constant mean and variance for the financial returns, namely a forecasting exercise that would not be possible were it not for particle filtering techniques. In section \textit{fill with the number of the section} we conclude by employing a non-parametric measure of realised volatility as a proxy for the latent volatility process, allowing a RMSE and MAE comparison between filters which is similar to that of section \textit{fill with the number of the section} for synthetic data.

@@

As regards the series that we will employ in section \textit{fill with the number of the section}, namely the series of realised volatilities which we will employ as a proxy for the true latent volatilities, the data were retrieved from the Oxford-Man Institute's Realized Library, which provides several different daily non-parametric measures of past volatility.\footnote{Such class of measures will be better presented in section \textit{fill with the number of the section}, when reviewing the use of realised volatility estimators as proxies for latent volatilities.} We consider a proper rescaling of the \textit{rv5} series, \textit{i.e.} model-free daily volatility estimates based on 5 min intraday return intervals. Also in this case, the chosen interval spanned from June 1st, 2017 to May 30th, 2021. \textit{Includere grafico e commento del grafico della realised volatility}

\subsection{Rescaling the data}

When specifying the parameters of a stochastic volatility model, one should clarify whether the series of observed financial returns is one of log returns or of percentage log returns. Indeed, let $y_t$ denote the returns at time $t$, so that a basic SV specification would prescribe $y_t|x_t\sim N(0,e^{x_t})$: clearly, for each period $t$, $e^{x_t}$ needs to be higher to replicate the observed \textit{percentage} log return rather than the observed log return. This implies that the latent volatility $x$ should fluctuate around higher values when using percentage log returns as $y_t$.
This feature, when adopting the aforementioned canonical SV model as defined in Kim et al. (1998), would be represented by a higher $\mu$. For example, as noted in Kastner (2019), daily log returns often have a variance of 0.0001 or less, which implies that $\mu$ should lie somewhere around $\ln(0.0001)\approx -9$; instead, when considering daily percentage log returns, one then has a variance of 1, so that $\mu=\ln(1)=0$. Note that when $\mu=0$, the canonical model boils down to the basic specification we employed on the simulated dataset, so that we prefer using percentage log returns for the 3 indices. In other terms, one might end up with the basic specification of the SV model when (roughly) calibrating the parameters of the canonical model on a wider set of financial returns.

Note that rescaling the log returns has analogous implications for the section in which we employ measures of realised volatility as a proxy for the latent volatility process.
Let us denote by $rv5_t$ the daily realised variance at time $t$, which we obtain from the \textit{rv5} series provided in the Oxford-Man database. Now, note that in a SV framework one should not directly compare $rv5_t$ to $x_t$, as $rv5_t$ can be rather seen as an estimate of the variance of the log returns at time $t$, \textit{i.e.} of $e^{x_t}$ in the SV model. Moreover, by construction of $rv5_t$,\footnote{Nota sulla formula, che considera returns, non returns in percentuale. Spiegare qui perchè possa avere senso paragonare questa varianza a quella dei log returns} the comparison should feature a SV model where the log returns did \textit{not} undergo the percentage transformation. Thus, in order to achieve a common scale with the percentage log returns, we need to multiply $rv5$ by $100^2$. To ease the interpretation and representation of the comparison between differently approximated filtering distributions, we finally take the square root of the rescaled $rv5$, so that we compare it to $e^{x_t/2}$ (\textit{In realtà detto così è poco preciso, perchè in realtà andiamo a confrontare con la trasformazione della media dello stato filtrato}).
